A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-09-27 paper/paper.tex
paper: MAJ

\section{Overture}
In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch14,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

\subsection{Augmented semi-simplicial sets}
We consider the category of augmented semi-simplices, which can be written in different ways, up to equivalence. Here, we describe the combinatorial definition.

\begin{notation}[Sequences of integers]
  We write $\{i_1, \ldots i_n\}$ for (ordered) sequences of integers. In particular, the empty sequence is written $\{\}$.
\end{notation}

\begin{definition}[\DeltaPlus]
  The category of augmented semi-simplices is defined in terms of objects and morphisms:

  \begin{align*}
    \obj(\DeltaPlus)   & := \Nat                                                                                    \\
    \hom(p, n)         & := \{i_{p + 1}, \ldots, i_n \mid 0 < i_{p + 1} < \ldots < i_n \leq n\}                     \\
    g \circ f          & := \{i_{j_{p + 1}}, \ldots, i_{j_n} \mid j_0 < i_{j_{p + 1}} < \ldots < i_{j_n} \leq j_n\} \\
    \id \in \hom(n, n) & := \{\}                                                                                    \\
  \end{align*}
\end{definition}

@@

  \begin{equation*}
    \Set_{\DeltaPlus} := \Set^{\op{\DeltaPlus}}
  \end{equation*}
\end{definition}

\begin{definition}[$\DeltaPlus^n$]
  The standard augmented $n$-semi-simplex $\DeltaPlus^n$ is defined as:
\end{definition}

We now illustrate the standard augmented $n$-semi-simplices for dimensions $1$, $2$, and $3$.

@@

  \begin{equation*}
    \begin{tikzcd}
      0
    \end{tikzcd}
  \end{equation*}
\end{example}

\begin{example}[$\DeltaPlus^2$]
  \begin{equation*}
    \begin{tikzcd}
      \kstar0 \arrow[r, dash, "\kstar\kstar"] & 0\kstar
    \end{tikzcd}
  \end{equation*}
\end{example}

@@

\begin{definition}[$\Cube^n$]
  The standard semi-cube $\Cube^n$ is defined as:
\end{definition}

We now illustrate the standard $n$-semi-cubes for dimensions $0$, $1$, and $2$.

\begin{example}[$\Cube^0$]
  $\Cube^0$ can be drawn as:

>>>>

\section{Overture}
In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch15,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

\subsection{Augmented semi-simplicial sets}
Augmented semi-simplicial sets are defined similarly to semi-simplicial sets, except that the points are additionally dependent on a ``color''. Conversely, semi-simplicial sets can be seen as augmented semi-simplicial sets over a singleton set of colors.

Let us associate dimension $0$ to colors; then, points are dimension $1$, lines are dimension $2$, and so on. There is hence a shift by one when compared to semi-simplicial sets. We can then draw augmented semi-simplicial sets like semi-simplicial sets, except for this shift by one.

While semi-simplicial sets are presheaves over the semi-simplex category $\hat{\Delta}$, augmented semi-simplicial sets are presheaves over, what we will define as $\DeltaPlus$. There are different ways to define $\DeltaPlus$, upto equivalence, and we use a combinatorial definition; the reason for our choice is that this definition has a straightforward extension to semi-cubical sets.

\begin{notation}[Sequences of integers]
  We write $\{i_1, \ldots i_n\}$ for sequences of integers. In particular, the empty sequence is written $\{\}$.

  Additionally, we define $i :: \{j_1, \ldots, j_n\}$ to be $\{i, j_1, \ldots, j_n\}$, and $\uparrow \{i_1, \ldots, i_n\}$ to be $\{i_1 + 1, \ldots, i_n + 1\}$.
\end{notation}

\begin{definition}[\DeltaPlus]
  The category of augmented semi-simplices is defined in terms of objects, morphisms, composition, and identity.

  \begin{align*}
    \obj(\DeltaPlus) & := \Nat                                                                    \\
    \hom(p, n)       & := \{\{i_{p + 1}, \ldots, i_n\} \mid 0 < i_{p + 1} < \ldots < i_n \leq n\} \\
    g \circ f        & :=
    \begin{cases}
      g                          & \text{if f = \{\}}                                    \\
      f                          & \text{if g = \{\}}                                    \\
      b :: (g \circ f')          & \text{if g = (a :: g'), f = (b :: f'), with b < a}    \\
      a :: (g' \circ \uparrow f) & \text{if g = (a :: g'), f = (b :: f'), with b \geq a} \\
    \end{cases}            \\
    \id              & := \{\}
  \end{align*}

  Composition can be shown to be associative.
\end{definition}

@@

  \begin{equation*}
    \Set_{\DeltaPlus} := \Set^{\DeltaPlus^{\mathsf{op}}}
  \end{equation*}
\end{definition}

\begin{definition}[$\DeltaPlus^n$]
  The standard augmented $n$-semi-simplex $\DeltaPlus^n$ is defined as:

  \begin{align*}
    \DeltaPlus^n    & : \Set_{\DeltaPlus}      \\
    \DeltaPlus^n(p) & := \Hom(p, n)            \\
    \DeltaPlus^n(f) & := \lambda g . g \circ f
  \end{align*}
\end{definition}

The standard $0$-semi-simplex is just the set of colors. Standard augmented $n$-semi-simplices have a geometric interpretation for dimensions $1$, $2$, and $3$, and we illustrate them now.

@@

  \begin{equation*}
    \begin{tikzcd}
      \star
    \end{tikzcd}
  \end{equation*}
\end{example}

\begin{example}[$\DeltaPlus^2$]
  In dimension $2$, it is
  \begin{equation*}
    \begin{tikzcd}
      \kstar0 \arrow[r, dash, "\kstar\kstar"] & 0\kstar
    \end{tikzcd}
  \end{equation*}
\end{example}

@@

\begin{definition}[$\Cube^n$]
  The standard semi-cube $\Cube^n$ is defined as:

  \begin{align*}
    \Cube^n    & : \CSet                  \\
    \Cube^n(p) & := \Hom(p, n)            \\
    \Cube^n(f) & := \lambda g . g \circ f
  \end{align*}
\end{definition}

Standard $n$-semi-cubes have a geometric interpreation, for dimensions $0$, $1$, and $2$, and we illustrate them now.

\begin{example}[$\Cube^0$]
  $\Cube^0$ is the empty set:
